---
title: "Evolution of hybrid legged robots with an open-loop controller"
output: html_notebook
---
```{r}
require(tidyverse)
```

## Load and clean evolution data

### Load
```{r}
d = read.csv2("../phase-1/results-evo.txt",dec=".")
summary(d)
```

### Clean
Extract key factors.
```{r}
d$controller = factor(str_match(d$target,"sin[PFA]+")[,1])
d$randomSeed = factor(str_match(d$randomGenerator,"=([0-9]+)")[,2])
d$nSoft = factor(str_count(d$target,"soft"))
d = d[,c("randomSeed","controller","nSoft","iterations","best.fitness.vx")]
summary(d)
```

## Evolution plot

```{r, fig.width=10, fig.height=4}
gd = d %>% group_by(controller, nSoft, iterations) %>% summarise_at(vars(best.fitness.vx), list(Q1=~quantile(.,probs=0.25), median=median, Q3=~quantile(.,probs=0.75)))
gd %>% ggplot(aes(iterations,median,col=nSoft)) + geom_line() + geom_ribbon(aes(ymin=Q1,ymax=Q3,fill=nSoft),alpha=0.2,color=NA) + facet_grid(.~controller) + theme_light() + scale_color_brewer(palette="Set1") + scale_fill_brewer(palette="Set1")
```

## Load and clean evolution data

## Load
```{r}
s = read.csv2("../phase-2/sections.txt",dec=".")
summary(s)
```

### Clean
Extract key factors.
```{r}
s$controller = factor(str_match(s$target,"sin[PFA]+")[,1])
s$randomSeed = factor(str_match(s$randomGenerator,"=([0-9]+)")[,2])
s$nSoft = factor(str_count(s$target,"soft"))
s = s[,c("randomSeed","controller","nSoft","iteration","dstIndex","stepIndex","d","pointQ")]
summary(s)
s = s %>% group_by(randomSeed, controller, nSoft, iteration, dstIndex) %>% mutate(srcQ = pointQ[1]) %>% ungroup() %>% mutate(rf=(pointQ-srcQ)/srcQ)
```

## Sections plots

### Actual fitness average across destinations, randomSeed 
```{r, fig.width=10, fig.height=4}
gs1 = s %>% group_by(controller, nSoft, iteration, stepIndex,d) %>% summarise_at(vars(pointQ), list(Q1=~quantile(.,probs=0.25), median=median, Q3=~quantile(.,probs=0.75)))
gs1 %>% ggplot(aes(d,median,col=nSoft)) + geom_line() + geom_ribbon(aes(ymin=Q1,ymax=Q3,fill=nSoft),alpha=0.2,color=NA) + facet_grid(factor(iteration)~controller) + theme_light() + scale_color_brewer(palette="Set1") + scale_fill_brewer(palette="Set1")
```

### Fitness relative variation average across destinations, randomSeed 
```{r, fig.width=10, fig.height=4}
gs2 = s %>% group_by(controller, nSoft, iteration, stepIndex, d) %>% summarise_at(vars(rf), list(Q1=~quantile(.,probs=0.25), median=median, Q3=~quantile(.,probs=0.75)))
gs2 %>% ggplot(aes(d,median,col=nSoft)) + geom_line() + geom_ribbon(aes(ymin=Q1,ymax=Q3,fill=nSoft),alpha=0.2,color=NA) + facet_grid(factor(iteration)~controller) + theme_light() + scale_color_brewer(palette="Set1") + scale_fill_brewer(palette="Set1") + ylim(c(-1,1))
```

### Fitness relative variation for iteration==100, randomSeed==0
```{r, fig.width=10, fig.height=8}
gs3 = s %>% filter(iteration==100 & randomSeed==0)
gs3 %>% ggplot(aes(d,rf,col=nSoft)) + geom_line() + facet_grid(factor(dstIndex)~controller) + theme_light() + scale_color_brewer(palette="Set1") + scale_fill_brewer(palette="Set1")
```

### Fitness relative variation for iteration==100
```{r, fig.width=10, fig.height=8}
gs4 = s %>% filter(iteration==100)
gs4 %>% ggplot(aes(d,rf,col=nSoft)) + geom_line(aes(linetype=controller)) + facet_grid(factor(dstIndex)~factor(randomSeed)) + theme_light() + scale_color_brewer(palette="Set1") + scale_fill_brewer(palette="Set1")
```

### Partial information count / number of changes (on relative fitness variation)
This function returns the ratio between the number (`which(cc)` below) of direction change in a sequence of values and the length of the sequence.
A variation is not considered if it is smaller than `epsilon`.
```{r}
changes = function(values, epsilon) {
  v = values[2:length(values)]-values[1:(length(values)-1)];
  v[abs(v)<=epsilon] = 0;
  c = sign(v)
  c = c[c!=0]
  cc = c[2:length(c)]!=c[1:(length(c)-1)]
  length(which(cc))/(length(values))
}
```

```{r, fig.width=10, fig.height=4}
s %>% group_by(controller, nSoft, iteration, randomSeed, dstIndex) %>% summarise_at(vars(rf), list(c10=~changes(.,0.10), c25=~changes(.,0.25), c50=~changes(.,0.50), c90=~changes(.,0.9))) %>% pivot_longer(c(c10,c25,c50,c90)) %>% group_by(controller,nSoft,iteration,name) %>% summarise_at(vars(value), list(Q1=~quantile(.,probs=0.25), median=median, Q3=~quantile(.,probs=0.75))) %>% ggplot(aes(x=iteration,y=median,color=nSoft)) + geom_line()  + geom_ribbon(aes(ymin=Q1,ymax=Q3,fill=nSoft),alpha=0.2,color=NA) + facet_grid(name~controller)  + theme_light() + scale_color_brewer(palette="Set1") + scale_fill_brewer(palette="Set1")
```
